---
output: html_document
editor_options: 
  chunk_output_type: inline
---

This R notebook contains the code used to generate a multi-resolution spatial mesh used for SpatioTemporal DeepKriging on Irish rainfall data.
The key step taken here is constrained Delaunay triangulation made available by the fdaPDE package (Arnore et al, 2025). 

```{r}

library(fdaPDE)
citation("fdaPDE")

```

First, we manually place knotpoints around the boundary of the region (Republic of Ireland). Here, we use a high density of knotpoints to emphasise any coastal effects.

```{r fig.height=12, fig.width=8}

# Boundary coordinates
boundary <- matrix(c(220, 350,   240, 410,   280, 430,   265, 470,   200, 455,
                     160, 425,   135, 385,   160, 360,   130, 355,   85, 360,
                     55, 335,    50, 300,    60, 290,    45, 270,    45, 250,
                     75, 210,    95, 190,    55, 130,    20, 105,    25, 70,
                     55, 35,     95, 15,     210, 55,    250, 85,    280, 90,
                     315, 95,    325, 120,   340, 160,   350, 190,   340, 235,
                     340, 255,   325, 280,   330, 310,   320, 325,   290, 325,
                     280, 355,   265, 360,   245, 340), 
                   nrow = 38, byrow = T)
segments <- cbind(1:nrow(boundary), c(2:nrow(boundary), 1))


# We'll make four meshes at increasing resolutions. create.mesh.2D 
# is from fdaPDE and constructs the mesh.

# First mesh
mesh_ireland_1 <- create.mesh.2D(boundary, 
                                 segments = segments,
                                 order = 1)

# refine.mesh.2D allows adjustments (we want to increase resolution)
mesh_ireland_1 = refine.mesh.2D(mesh_ireland_1,
                                minimum_angle = 35)

# To make the next mesh, we can adjust the previous one
mesh_ireland_2 = refine.mesh.2D(mesh_ireland_1,
                                minimum_angle = 25,
                                maximum_area = 400)

# Third mesh
mesh_ireland_3 = refine.mesh.2D(mesh_ireland_2,
                                minimum_angle = 25,
                                maximum_area = 100)

# Fourth mesh
mesh_ireland_4 = refine.mesh.2D(mesh_ireland_3,
                                minimum_angle = 30,
                                maximum_area = 50)

# Here we adjust the coordinates to match easting and 
# northing used for the Irish grid
df1 <- mesh_ireland_1$nodes * 1000
df2 <- mesh_ireland_2$nodes * 1000
df3 <- mesh_ireland_3$nodes * 1000
df4 <- mesh_ireland_4$nodes * 1000

names(df1) <- c("east", "north")
names(df2) <- c("east", "north")
names(df3) <- c("east", "north")
names(df4) <- c("east", "north")

# write.csv(df1, "data/mesh_1.csv")
# write.csv(df2, "data/mesh_2.csv")
# write.csv(df3, "data/mesh_3.csv")
# write.csv(df4, "data/mesh_4.csv")

par(mfrow = c(2, 2))
plot(mesh_ireland_1, main = paste0("Mesh 1: ", 
                                   nrow(mesh_ireland_1$nodes), " nodes"))
plot(mesh_ireland_2, main = paste0("Mesh 2: ", 
                                   nrow(mesh_ireland_2$nodes), " nodes"))
plot(mesh_ireland_3, main = paste0("Mesh 3: ", 
                                   nrow(mesh_ireland_3$nodes), " nodes"))
plot(mesh_ireland_4, main = paste0("Mesh 4: ", 
                                   nrow(mesh_ireland_4$nodes), " nodes"))

```
